Equity conceptualization and standards within transportation literature: a scoping review

### 2023-09-28 Manuscript/Manuscript.qmd
Antonio is reviewing and editing Manuscript.qmd

>>>>

On July 20, 2021, at 9:12 a.m. (EDT), a rocket took off from a launching pad in West Texas. On board was a small group of four passengers that included Jeff Bezos, then the world's richest person [@harwoodBezos2021]. The flight that morning was one of the first crewed private suborbital flights, and the adventure (described as "intense" by a member of the crew) lasted a total of 10 minutes and 10 seconds [@harwoodBezos2021]. In addition to intense, it was expensive: a seat for the flight had previously been auctioned for no less than $28 million USD [@griffinPerson2021]. It was not yet 10:00 am (EDT) after the rocket's return from its excursion to the very edge of atmosphere, when Bezos declared that this was the "[b]est day ever". To reporters covering the event, he said "I want to thank every Amazon employee and every Amazon customer, because you guys paid for all of this" [@johnsonRoadToSpace2021].

Meanwhile, back on planet Earth, the employees that Bezos thanked for paying for his suborbital escapade were facing some very mundane problems of their own, and none as lofty as conflicting schedules that prevented them from flying in rockets [@griffinPerson2021]. According to reports, people employed directly or indirectly by Amazon for warehouse or delivery work had, for years, been treated to "inhumane" conditions [@fungEuropean2018; @scottAmazon2019; @greeneAlgorithm2021], and subjected to surveillance on the job, crushing demands for productivity quotas that led to injury, enjoying little or no job security, and more generally treated as disposable inputs to Amazon's efficiency goals and consumers' desires [@tungAmazon2020; @reeseTeamsters2022; @middletonRobot2023].

Coverage by the mainstream media of the July 20 launch was in many cases uncritical. "We're going to build a road to space so that our kids and their kids can build a future," Bezos declared, and added "...we need to do that to solve the problems here on Earth" [@johnsonRoadToSpace2021]. Few reporters thought to ask questions about what problems the billionaire planned to solve on Earth, or what kind of future Bezos was trying to build, and for whose children. A less dispassionate observer might be excused for wondering about the basic fairness of a man amassing a nearly unimaginable fortune that allowed him to build and fly his own spaceship, while masses of his employees were treated as throwaway cogs in the vast apparatus of his logistics empire.

But the question of fairness is not a simple one, and certainly not one amenable to scientific inquiry. What constitutes "fair" would likely look very different in democratic societies, compared to societies where no one owns anything—perhaps not even their own time or bodily autonomy—because everything belongs to some collective illusion (e.g., the state, a monarch). As well, what is fair would look very different in a society where very few own most, and most own very little due to a different illusion (e.g., that wealth equates merit). In these two cases, it is not necessarily difficult to elucidate the meaning of fair: fair is what the state, the monarch, or the extremely wealthy say it is. In contrast, in democratic societies (and democracy itself is a collective illusion) individuals are accorded rights, and there the relevant question is what a fair distribution of the burdens and benefits of things, from roads to space travel, would look like.

Rockets are transportation technologies, tools of space-time convergence. By enabling movement at very high speeds they might prove important to the expansion of the human species beyond our home planet—at some indefinite time in the future. But in the present moment, the benefits of a private suborbital flight (e.g., the joy of movement, the sense of adventure, the awe of seeing Earth from space), accrue to just a few, whereas the burdens (e.g., the climate-altering emissions) affect all, and not even equally at that. In a democratic society, the members of the collective might decide that this arrangement, whereby large rewards are offered to the most qualified individuals (e.g., Bezos) to entice them to take more important positions [e.g., creating and leading Amazon; see @davisSome1945], is in fact a fair distribution of the burdens and benefits of space travel. Then again, the members of the collective might decide otherwise.

Multiple studies in an international context suggest that most people in many societies have some tolerance for inequality [@kiatpongsanHow2014]: indeed, it appears that _some_ stratification, as suggested by @davisSome1945, is perceived as being functional. However, extreme inequality is often frowned upon, and can lead to social dysfunction [@acemogluWhy2000; @taydasCan2012; @duIncome2019; @houleInequality2022]. But the perceptions of what is "fair" are malleable, and can be affected by opportunities for social mobility [@shariffIncome2016], exposure to inequality [@schroderIncome2017; @garciaChanging2023], **learned helplessness**, etc.

Positive -> what is the current status of the distribution of benefits and burdens?
Normative -> what should it be?

### 2023-09-28 Manuscript/Manuscript_old.Rmd
Antonio is reviewing and editing Manuscript.qmd

---
title: Equity conceptualization and standards within transportation literature: a scoping review
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.

>>>>

---
title: 'Equity conceptualization and standards within transportation literature: a scoping review'
author:
  - name: Alice Anonymous
    email: alice@example.com
    affiliation: Some Institute of Technology
    footnote: 1
  - name: Bob Security
    email: bob@example.com
    affiliation: Another University
  - name: Cat Memes
    email: cat@example.com
    affiliation: Another University
    footnote: 2
  - name: Derek Zoolander
    email: derek@example.com
    affiliation: Some Institute of Technology
    footnote: 2
address:
  - code: Some Institute of Technology
    address: Department, Street, City, State, Zip
  - code: Another University
    address: Department, Street, City, State, Zip
footnote:
  - code: 1
    text: "Corresponding Author"
  - code: 2
    text: "Equal contribution"
abstract: |
  This is the abstract.
